---
title: "PSTAT 10 Homework 1"
subtitle: "Due 8/10/22 11:59pm"
author: "Ingmar Sturm"
date: "`r Sys.Date()`"
output: pdf_document
---

Hand in your homework on time! Your TA has a life and will want to grade your
homework shortly after it is due. Late homework assignments will receive a 
penalty.

# Problem 1: Exploring the ecosystem

Here is a list of five R packages
* beepr
* fun
* fortunes
* cowsay
* praise

Choose one of these packages and download its source code. The list of all packages on CRAN is here
https://mirror.las.iastate.edu/CRAN/. On a package’s page, you can find the *package source* to download,
which will be in tar.gz format. This is called a _tarball_.

Unzip the tarball, navigate to an R script, and copy and paste a function’s name and arguments (but not the
body) to your homework solution. Comment on the name of the function and how many arguments it has.
For example in cowsay, the script in `cowsay/R/utils.R` contains a function `check_color`:

```{r, eval=FALSE}
check_color <- function(clr) {
```

It takes one argument, `clr`.
If you’re wondering how I included an incomplete piece of code in my R Markdown, it is because I added
`eval = F` to my code chunk, like this: `{r eval=F}`.

# Problem 2: More ecosystem exploration

The point of this problem is to demonstrate that a community of developers are constantly improving the R
language.

If you installed R recently, you should be on version 4.2.0 (to check, run R.Version() in the console).

Describe one feature or bug fix that is new in version 4.2.0. This can be found via the *what’s new* link in
https://mirror.las.iastate.edu/CRAN/. You can simply copy and paste the description for your solution to
this problem.

# Problem 3: State areas

Load the library called datasets. The vectors state.area and state.name contain in alphabetical order
the land area (in square miles) and the names of each state in America.

1. Find the mean area of all states.
2. Find the median area of all states.
3. Find the name and the area of the smallest state.
4. Find the name and the area of the largest state.

There are many ways to do parts 3 and 4, but one way is to use which.min and which.max.

# Problem 4: A function without parameters

Most of functions you have encountered so far have parameters. It is also 
possible to create and call functions without parameters, such as `Sys.time()`
or `getwd()`. This can be useful if you are writing complex code such as an 
R package and you want to reduce the amount of code you have to handle.

Note: A parameter is a variable in a function definition. It is a placeholder and 
hence does not have a concrete value. An argument is a value passed during 
function invocation. In a way, arguments fill in the place the parameters have 
held for them.

Write a function `my_time` that takes no arguments and prints `Hello, the time
is` followed by the actual current time (but not the date), all in one sentence. 
Take a look at the help files of the `paste0`, `Sys.time`, and `strptime` 
functions if you get stuck.

```{r include=FALSE}
my_time <- function(){
  # any time format is fine
  paste0("Hello, the time is ", format(Sys.time(), "%H:%M"), ".") 
}
```


# Problem 4: Length of a vector

The length of a vector is the square root of the sum of the squares of the 
horizontal and vertical components. Write a function that takes a vector as its
input and returns the length of the vector as its output. If the the vector is
not numeric, print the message `The argument must be a numeric vector!`

Test the function with these examples

```{r include=FALSE}
vector_length <- function(x){
  if(!is.numeric(x)){
    print("The argument must be a numeric vector!")
  } else {
    my_vec_length <- sqrt(sum(x^2))
    return(my_vec_length)
  }
}
```

```{r}
vector_length(c(5, 9, 11, 15))
```

```{r}
vector_length(c("mouse"))
```

# Problem 4: Dot product

The dot product of two vectors $(x_1, x_2, \dots , x_n)$ and $(y_1, y_2, \dots , y_n)$ is the number

$$
\sum^{n}_{t=1}=x_iy_i=x_1 \times y_1 + x_2 \times y_2 + \dots + x_n \times y_n.
$$

Write a function dot_product that takes two numeric vectors and returns their dot product. If either
argument is not numeric, print the message `Both arguments must be numeric!`

```{r include=FALSE}
dot_product <- function(x, y) {
  if(!is.numeric(x) | !is.numeric(y)){
    print("Both arguments must be numeric!")
  } else {
  my_dot_prod <- sum(x*y)
  return(my_dot_prod)
  }
}
```

Test your function on the following inputs:

```{r eval=FALSE}
dot_product(1:3, c(0, 1, 5))
```

```{r eval=FALSE}
dot_product(2, 4)
```

```{r eval=FALSE}
dot_product(c(1, 1), c("dog", "cat"))
```

# Problem 5: Frobenius norm

The Frobenius norm of a matrix is the square root of the sum of the squares of all its entries. For example,
the Frobenius norm of the matrix 

$$
\begin{bmatrix}
1 &2\\
3 &4
\end{bmatrix}
$$

is $\sqrt{1^2 + 2^2 + 3^2 + 4^2} = \sqrt{1 + 4 + 9 + 16} = \sqrt{30} \approx 5.48$. Write a function frobenius_norm that takes
a numeric matrix and returns its Frobenius norm. If the input is not numeric print Argument must be
numeric! and if the input is not a matrix, print `Argument must be a matrix!.`

```{r include=FALSE}
frobenius_norm <- function(x){
  if(!is.matrix(x)){
    print("Argument must be a matrix!")
  } else if(!is.numeric(x)){
    print("Argument must be numeric!")
  } else {
  my_norm <- sqrt(sum(x^2))
  return(my_norm)
  }
}
```

Test your function on the following inputs:
```{r}
frobenius_norm(matrix(1:4, nrow = 2, ncol = 2))
```

```{r}
frobenius_norm(c(3,5,7,10,15,21))
```

```{r}
frobenius_norm(matrix(c(3,5,7,10,15,21), nrow = 2, ncol = 3))
```

```{r}
frobenius_norm(matrix(c(3,"fish",7,10,15,21), nrow = 2, ncol = 3))
```

# Problem 6: Compare Count

Write a function compare_count that takes three arguments: x, y, comp. Arguments x and y are numeric
vectors of the same length, and comp is one of the three characters (strings) >, <, =. The function
compare_count returns the number of elements in x that are greater than, less than, or equal to the
corresponding element in y, depending on the value of comp. The comp argument should default to “>”. If
these assumptions are not met print error messages that match the sample output below. If there is an error,
do not do any further operations.

```{r, eval=FALSE}
compare_count(rep(1, 5), rep(2, 5))
```

```
## [1] 0
```

```{r, eval=FALSE}
compare_count(rep(1, 5), rep(2, 5), ">")
```

```
## [1] 0
```

```{r, eval=FALSE}
compare_count(c(1, 2, 1, 2, 1), rep(2, 5), "<")
```

```
## [1] 3
```

```{r, eval=FALSE}
compare_count(c(1, 2, 1, 2, 1), rep(2, 5), "=")
```

```
## [1] 2
```

```{r, eval=FALSE}
compare_count(c(1, 2, 1, 2, 1), rep(2, 5), ">=")
```

```
## [1] "Unrecognized compare operator!"
## NULL
```

```{r, eval=FALSE}
compare_count(c(1, 2, 1, 2, 1), rep(2, 6), "=")
```

```
## [1] "Both vectors must have the same length!"
## NULL
```

```{r, eval=FALSE}
compare_count(c(1, 2, 1, 2, "owl"), rep(2, 6))
```

```
## [1] "Both vectors must be numeric!"
## NULL
```

# Problem 7: Styling your code

Writing readable code is important and it is best to get it right the first time 
around. This will not only help your collaborators read your code but also
yourself.

Read [Chapter 2: Syntax](https://style.tidyverse.org/syntax.html) of the
tidyverse style guide. Then go through your code and make sure it conforms to
the tidyverse style guide.







