# Lecture 2 demo code

#### 2022-06-24

### Filtering

1.  

```{r}
gas_bill <- c(46, 33, 39, 37, 46, 30, 48, 32, 49, 35, 30, 48)
length(gas_bill)
```

2.  

```{r}
sum(gas_bill > 40)
```

The expression `gas_bill > 40` is a logical vector. Summing this vector
gives the number of TRUE elements.

3.  

```{r}
mean(gas_bill > 40) * 100
```

The mean of a logical vector is the proportion of TRUE elements.

4.  

```{r}
sum(gas_bill[gas_bill > 40])
```

### Recycling (slide 8)

1.  

```{r}
gas_bill_update <- gas_bill + c(-3, 7)
```

The shorter vector `c(-3, 7)` is recycled to match the length of
`gas_bill`.

2.  

```{r}
sum(gas_bill - gas_bill_update)
```

So I owe the company 24 dollars.

### Vectorization (slide 11)

1.  

```{r}
sum(sqrt(1:100))
```

2.  

```{r}
prod(log(100:200))
```

3.  

```{r}
ints <- 1:100
ints_sq <- ints^2
sum(ints[ints_sq > 300 & ints_sq < 500])
```

Pay special attention to the expression `ints_sq > 300 & ints_sq < 500`
above.

### Matrices (slide 18)

1.  

```{r}
(mx <- matrix(1:17, nrow = 4, ncol = 5))
```

The first three entries of `1:17` were recycled. The warning tells us
the recycling was not complete.

2.  

```{r}
mx[c(1, 4), c(1, 2, 5)]
```

3.  

```{r}
mx[mx > 10] <- 0
mx
```

First, `mx[mx > 10]` filters for the entries larger than 10. Then, `0`
is recycled to match the length of these entries. Finally, the entries
are each assigned to 0.
