# Lecture 4 demo code

#### 2022-06-27

## Remove element (slide 4)

### Loop approach

Let's try using a loop first. A preliminary idea is as follows: 1.
Construct a result vector `result` 2. Loop through the input `v` - if
`v[i]` does not equal `target`, add it to `result` - otherwise, do not
add it to `result` (do nothing) 3. Return `result`

Here is this idea implemented:

```{r}
remove_elt <- function(v, target) {
  result <- numeric(0)
  for (i in seq_along(v)) {
    if(v[i] != target) {
      result <- c(result, v[i])
    }
  }
  return(result)
}
remove_elt(c(14, 14, 7, 7, 14, 10), 14)
```

### Vectorized approach

Instead of a loop, we can do a *vectorized* comparison and then filter
for the desired values of `v`.

```{r}
remove_elt <- function(v, target) {
  v[v != target]
}
remove_elt(c(14, 14, 7, 7, 14, 10), 14)
```

In the above, `v != target` is a vector of logicals which we use to
subset `v` itself. Please make sure you are comfortable with this
operation.

## Print Square (slide 5)

### Matrix approach

The first approach is clever: just print a matrix.

```{r}
print_square <- function(n) {
  m <- matrix("*", nrow = n, ncol = n)
  print(m)
}

print_square(5)
```

Think about how *recycling* occurs so that a single `*` becomes `n^2`
`*`s.

### Loop approach

Let's build a vector of `n` `*`s and then use a loop to print it `n`
times.

```{r}
print_square <- function(n) {
  asts <- rep("*", n)
  for (i in seq_len(n)) {
    print(asts)
  }
}
print_square(4)
```

## Print triangle (slide 6)

A matrix approach is less straightforward in this case. But we can
easily adapt the above loop. Now, the number of asterisks depends on the
loop index.

```{r}
print_triangle <- function(n) {
  for (i in seq_len(n)) {
    asts <- rep("*", i)
    print(asts)
  }
}
print_triangle(4)
```

## Max vector (slide 7)

### Loop with comparison

We want to compare `nums1` and `nums2` element-wise, so it makes sense
to loop over them. The vectors are guaranteed to be of length `n`, so we
can use this.

```{r}
max_vec <- function(nums1, nums2, n) {
  result <- vector(length = n) # this initializes an "empty" vector with length n
  for (i in seq_len(n)) {
    if (nums1[i] < nums2[i]) {
      result[i] <- nums2[i] # Since nums2[i] is larger, place it into the result
    } else {
      result[i] <- nums1[i] # Since nums1[i] is larger, place it into the result
    }
    # Note that if they are equal, it does not matter which we put in.
  }
  return(result)
}
max_vec(1:4, c(0, 27, -5, 19), 4)
```

### Loop with builtin max

It is always good to check if R already implements some basic operation.
In this case, `max(a, b)` returns the larger of `a` and `b`.

```{r}
max_vec <- function(nums1, nums2, n) {
  result <- vector(length = n)
  for (i in seq_len(n)) {
    result[i] <- max(nums1[i], nums2[i]) # use the builtin max function
  }
  return(result)
}
max_vec(1:4, c(0, 27, -5, 19), 4)
```

The above saves us from explicitly writing a branch.

### pmax

A *even better* way is to use Google to find the `pmax` function, which
does everything for us.

```{r}
pmax(1:4, c(0, 27, -5, 19))
```

## Small Count Simple (slide 8)

It is possible to do this with a loop. But let's use a vectorized
approach:

```{r}
small_count_simple <- function(v, target) {
  sum(v < target)
}
small_count_simple(c(10, 15, -2, 5), 11)
```

`v < target` results in a logical vector. Summing up the result gives
the number of TRUE elements.

## Small Count (slide 9)

Loop over a `result` vector and fill in the values using the approach
from the previous problem.

```{r}
small_count <- function(v) {
  result <- vector(length = length(v))
  for (i in seq_along(v)) {
    result[i] <- sum(v < v[i])
  }
  return(result)
  small_count(c(12, 100, -3, -12))
}
```

## Two Sum (slide 10)

For this problem, we must be able to enumerate pairs of indices going
from `1` to `length(v)`. In other words, if `v` has 4 elements, we need
to check the following pairs:

> (i, j) \<= (1, 2), (1, 3), (1, 4), (2, 3), (2, 4), (3, 4)

Then if we look at `v[i] + v[j]`, we will see the sum of all pairs of
elements of `v`. Think: why were `(2, 1), (4, 2), ...` not included?

### Loop approach

The idea is to use two loops: `i` going from 1 to `length(v)` and `j`
going from 1 to `i-1`. This ensures we go over each pair exactly once
(why?).

```{r}
two_sum <- function(v, target) {
  i <- 0
  j <- 0
  for (i in seq_along(v)) {
    for (j in seq_len(i-1)) {
      if (v[i] + v[j] == target) {
        return(c(i, j))
      }
    }
  }
}
two_sum(c(2, 7, 11, 15), 9)
```

## First Even (slide 11)

The expression `(v %% 2) == 0` gives a vector of logicals indicating
TRUE if the entry of `v` is even and FALSE otherwise. Calling `which` on
this logical vector gives all the indices of TRUE, which are all the
indices of even numbers of `v`. Then index with `[1]` to find the first
such index.

```{r}
first_even <- function(v) {
  w_idx <- which((v %% 2) == 0)[1]
  v[w_idx]
}
first_even(c(-1, -4, 0, 2))
```

It may be clearer to do the indexing *after* subsetting `v`: Below,
`w_idx` is now a *vector* of indices, which we use to subset `v`.

```{r}
first_even <- function(v) {
  w_idx <- which((v %% 2) == 0)
  (v[w_idx])[1]
}
first_even(c(-1, -4, 0, 2))
```

## Last Negative (slide 12)

This is an **exercise**; please do this on your own. If you are stuck,
talk to me.

## Any and all (slide 13)

This is also an **exercise**; please do this on your own. If you are
stuck, talk to me.
