# Lecture 7 demo code

#### Ingmar

#### 2022-06-30

We will need the following packages:

```{r}
library(tidyverse)
```

```{r}
library(gapminder)
library(datasets)
```

## Create a scatter plot (slide 7)

```{r}
plot(airquality$Solar.R, airquality$Ozone)
trend <- lm(airquality$Ozone ~ airquality$Solar.R)
abline(trend, col = "red")
```

## Layering plots demo 1 (slide 17)

```{r}
plot(airquality$Wind, airquality$Ozone, type="n", 
     main = "Ozone and Wind in NYC",
     xlab = "Wind", ylab = "Ozone")

notMay <- filter(airquality, Month != 5)
may <- filter(airquality, Month == 5)

points(notMay$Wind, notMay$Ozone, col = "red", pch = 19)
points(may$Wind, may$Ozone, col = "blue", pch = 17)
legend("topright", pch = c(17, 19), col = c("blue", "red"),
       legend = c("May", "Other months"))
```

## Layering plots demo 2 (slide 18)

First is the *data munging* step: manipulating the data frame so we have
the necessary variables.

Then we use those variables in a plot.

```{r}
# Data munging
airqual_hc <- airquality |>
                mutate(hotorcold =
                         as.factor(
                           ifelse(Temp > median(Temp), "hot", "cold"))) |>
                select(Ozone, Wind, hotorcold)
airqual_hot <- airqual_hc |>
                  filter(hotorcold == "hot")
airqual_cold <- airqual_hc |>
                  filter(hotorcold == "cold")

# Plotting
plot(airquality$Wind, airquality$Ozone, type = "n",
     main = "Ozone and Wind in NYC, 1973",
     xlab = "Wind (mph)", ylab = "Ozone (ppb)")
points(airqual_hot$Wind, airqual_hot$Ozone, pch = 19, col = "tomato")
points(airqual_cold$Wind, airqual_cold$Ozone, pch = 19, col = "cadetblue3")
trend <- lm(airquality$Ozone ~ airquality$Wind)
abline(trend)
legend("topright", pch = 19, col = c("tomato", "cadetblue3"),
       legend=c("Hot", "Cold"))
```

## Plot of powers (slide 23)

```{r}
x <- seq(-1, 5, length = 200)
y2 <- x^2
y3 <- x^3
plot(y2 ~ x, type = "l", col = "tomato", lty = 2, lwd = 2,
     ylab = "y", main = "Plot of Power")
lines(y3 ~ x, type = "l", col = "blue", lwd = 2)
legend("topright", lwd = 2, lty = c(2, 1), col = c("tomato", "blue"), 
       legend = c("x^2", "x^3"))
```

### What happens if we plot the cubic curve first?

```{r}
x <- seq(-1, 5, length = 200)
y2 <- x^2
y3 <- x^3
plot(y3 ~ x, type = "l", col = "blue", lty = 1, lwd = 2,
     ylab = "y", main = "Plot of Powers")
lines(y2 ~ x, type = "l", col = "tomato", lwd = 2, lty = 2)
legend("topright", lwd = 2, lty = c(2, 1), col = c("tomato", "blue"), 
       legend = c("x^2", "x^3"))
```

Why is this result different? Because `y3`, the vector containing the
cubes, spans a larger range than `y2`, the vector containing squares.
The initial call to `plot` sets the boundary depending on this range.
