---
title: "PSTAT 10 Homework 1 Solutions"
output: html_document
---

# Problem 1: Exploring the ecosystem

(Unzip the tarball, navigate to an R script, and copy and paste a function’s 
name and arguments (but not the body) to your homework solution. Comment on the 
name of the function and how many arguments it has.)

Various answers.

# Problem 2: More ecosystem exploration

CHANGES IN R 4.4.1:

* C-LEVEL FACILITIES:

    - Functions R_atof and R_strtod declared in header R_ext/Utils.h
      are now documented in 'Writing R Extensions' and so formally part
      of the API.

    - The non-API entry points Rf_setSVector, Rf_StringFalse,
      Rf_StringTrue and Rf_isBlankString have been added to those
      reported by R CMD check.

    - The new function Rf_allocLang is now available. This provides an
      alternative to the idiom of calling Rf_allocList followed by
      SET_TYPEOF.

*  UTILITIES:
    - R CMD check now reports as warnings what gfortran calls 'Fortran
      2018 deleted features', all of which have long been marked as
      'obsolescent' and some of which were deleted in Fortran 2008 or
      earlier.  Fortran compilers are no longer required to support
      these.

*  BUG FIXES:
    - as.numeric(), scan(), type.convert() and other places which use
      the internal C function R_strtod now require a _non-empty_ digit
      sequence in a decimal or binary exponent.  This aligns with the
      C/POSIX standard for strtod and with ?NumericConstants.
    - as.data.frame(m, make.names=NA) now works correctly for a matrix
      m with NA's in row names.
    - The error message from <POSIXlt>[["hour"]] and similar now
      mentions *[[, "hour"]], as wished for in PR#17409 and proposed by
      Michael Chirico.
    - qbinom() and potentially qpois(), qnbinom(), no longer sometimes
      fail accurate inversion (of pbinom(), etc), thanks to Christopher
      Chang's report and patch in PR#18711.
    - The internal help server on Windows can again serve requests sent
      in quick succession, fixing a regression in R 4.4.0.
    - debugcall(<S3Generic>()) now also works when a corresponding
      S4-generic version is in the methods cache (PR#18143).
    - Package tools' toTitleCase(ch0) now returns character(0) when ch0
      is of zero length; fixing PR#18724, reported by David Hugh Jones.
    - R CMD check is no longer broken (without a check result and no
      explanation in 00check.log) for a package which declares an
      invalid VignetteBuilder in DESCRIPTION but has no vignettes.

# Problem 3: State areas

```{r}
library(datasets)
```


Find the mean area of all states.

```{r}
mean(state.area)
```


Find the median area of all states.

```{r}
median(state.area)
```

Find the name and the area of the smallest state.

```{r}
smallest <- which.min(state.area)
c(state.area[smallest], state.name[smallest])
```

Find the name and the area of the largest state.
```{r}
largest <- which.max(state.area)
c(state.area[largest], state.name[largest])
```

# Problem 4: Dot product

```{r}
dot_product <- function(x, y) {
  if (is.numeric(x) && is.numeric(y)) {
    return(sum(x * y))
  }
  print("Both arguments must be numeric!")
}
dot_product(1:3, c(0, 1, 5))
```

```{r}
dot_product(2, 4)
```

```{r}
dot_product(c(1, 1), c("dog", "cat"))
```


# Problem 5: Frobenius norm
```{r}
frobenius_norm <- function(mx) {
  if (!is.numeric(mx)) {
    print("Argument must be numeric!")
  } else if (!is.matrix(mx)) {
    print("Argument must be a matrix!")
  } else {
    sqrt(sum(mx^2))
  }
}
frobenius_norm(matrix(1:4, nrow = 2, ncol = 2))
```

```{r}
frobenius_norm(c(3,5,7,10,15,21))
```

```{r}
frobenius_norm(matrix(c(3,5,7,10,15,21), nrow = 2, ncol = 3))
```

```{r}
frobenius_norm(matrix(c(3,"fish",7,10,15,21), nrow = 2, ncol = 3))
```

# Problem 6: Compare Count

```{r}
compare_count <- function(x, y, comp = ">") {
  if(!is.numeric(x) || !is.numeric(y)) {
    print("Both vectors must be numeric!")
    return()
  }
  if (length(x) != length(y)){
    print("Both vectors must have the same length!")
    return()
  }
  
  if (!(comp %in% c('>', '<', '='))) {
    print("Unrecognized compare operator!")
    return()
  }
  
  if (comp == '>') {
    return(sum(x > y))
  }
  
  if (comp == '<') {
    return(sum(x < y))
  }
  
  if (comp == '=') {
    return(sum(x == y))
  }
}
compare_count(rep(1, 5), rep(2, 5))
```


```{r}
compare_count(rep(1, 5), rep(2, 5), ">")
```


```{r}
compare_count(c(1, 2, 1, 2, 1), rep(2, 5), "<")
```

```{r}
compare_count(c(1, 2, 1, 2, 1), rep(2, 5), "=")
```

```{r}
compare_count(c(1, 2, 1, 2, 1), rep(2, 5), ">=")
```

```{r}
compare_count(c(1, 2, 1, 2, 1), rep(2, 6), "=")
```

```{r}
compare_count(c(1, 2, 1, 2, "owl"), rep(2, 6))
```

# Problem 7: Styling your code

Code should be styled according to tidyverse style guide.
