---
title: PSTAT 10 Homework 2
subtitle: Due 7/8/24 11:59pm
output:
  pdf_document:
    latex_engine: xelatex

---

# Problem 1: Convergence to e

The number $e \approx 2.718$ can be expressed in many different ways. One way is as a limit and another is as an infinite series:

$$
e = \lim_{n \to \infty} \left( 1 + \frac{1}{n} \right)^n \quad \text{limit representation}
$$

$$
e = \sum_{k=0}^{\infty} \frac{1}{k!} \quad \text{series representation}
$$
Create two vectors containing these values at the points $x = 1, 2, . . . , 100$. Then create the following plot in
base R to assess the convergence of these two representations.


```{r fig.align = "center", fig.dim=c(6, 4.5), out.width="70%", out.height="70%", echo=FALSE}

x <- 1:100


limit <- (1 + 1/x)^x
series <- sapply(x, function(n) sum(1/factorial(0:n)))


plot(x, limit, type = "l", col = "blue", xlab = "x", ylab = "y", main = "Convergence to e")
lines(x, series, col = "red")


legend("bottomright", legend = c("Limit", "Series"), col = c("blue", "red"), lty = 1)

```
*Hint*: One way is to start with the following:

```r
x <- 1:100
e_limit <- vector(length = length(x))
e_limit[1] <- 1
e_series <- vector(length = length(x))
e_series[1] <- 1
```
Then you can use a **for loop** to fill in the values.

# Problem 2: nycflights13

For this problem, we will work with the `flights` data set in the nycflights13 package. Install and load the `nycflights13` package. Also make sure you load `tidyverse` in case you need it.

``` r
library(nycflights13)

library(tidyverse)
```

1.  Provide a brief description of the data set and a few of the variables (use `?flights`). Is `flights` a tibble?

2.  Extract a tibble containing American Airlines (AA) flights to LAX that departed before 1030. Return only the columns `month`, `day`, `dep_time`, `dest`, and `carrier`. How many flights fit these criteria?

3. Extract a tibble containing all flights on Christmas day; 12/25/2013. **What is the total number of
miles** traveled across all flights on this day?

4. The air_time variable gives the duration of the flight in minutes. Create a tibble containing flights on
Christmas day and only the variables month, day, origin, dest, and air_time_hour where the last
variable gives the duration of the flight in *hours*.

```
## A tibble: 719 x 5
## month   day  origin  dest  air_time_hour
## <int> <int> <chr>  <chr>        <dbl>
1    12    25   EWR   CLT           1.63
2    12    25   EWR   IAH           3.38
3    12    25   JFK   MIA           2.43
4    12    25   JFK   BQN           3.18
5    12    25   LGA   ORD           2.05
6    12    25   LGA   DTW           1.47
7    12    25   LGA   ATL           1.97
8    12    25   LGA   FLL           2.45
9    12    25   EWR   FLL           2.48
10   12    25   JFK   MCO           2.28
# . . . with 709 more rows
```

# Problem 3: mtcars

For this problem we will work with the `mtcars` data set in the `datasets` library.

```{r}
library(datasets)
```
If you’re not happy with the size and scale of your plots, these can be changed via RMarkdown options. For example, my barplot in part 4 of this problem uses the following options: 

{r, fig.align = "center", fig.dim=c(4.5, 4.5), out.width="40%", out.height="40%"}

fig.dim specifies the dimensions of the figure while the out parameters are used to scale the figure.

1.  Provide a brief description of this data set and some of the variables. Is it a tibble?

2.  Create the following histogram of displacement along with vertical lines indicating the mean and median displacement. *Hint:* I set the breaks to go from 0 to 500 in increments of 25.

``` {r fig.align = "center", fig.dim=c(5, 4.5), out.width="40%", out.height="40%", echo=FALSE}
hist(mtcars$disp, breaks=seq(0, 500, 25),
     main= "Hist of Displacement",
     xlab = "Displacement (cu.in.)")
abline(v = c(mean(mtcars$disp), median(mtcars$disp)),
       lty=c(2,3), lwd=2, col="red")
legend("topright", legend = c("mean disp", "median disp"),
       lty=c(2,3), lwd=2, col="red")
```

3.  Create the following boxplot of the 1/4 mile time against the number of cylinders. There is one outlier. Which car is the outlier? To answer this, note that the names of the cars are row names in the data set.

``` {r fig.align = "center", fig.dim=c(5, 4.5), out.width="40%", out.height="40%", echo=FALSE}
boxplot(mtcars$qsec ~ mtcars$cyl,
        xlab = "Number of Cylinders", ylab="1/4 mile time (sec)")
```

4. Create the following barplot of counts of cars with different numbers of cylinders.

``` {r fig.align = "center", fig.dim=c(4.5, 4.5), out.width="40%", out.height="40%", echo=FALSE}
barplot(table(mtcars$cyl), ylab="Count", xlab="Number of cylinders")
```


# Problem 4: Search insert position

Write a `function search_insert_position(v, target)` which takes a sorted numerical vector `v`, a numerical target `target`. If target is present in `v`, return the index of `target`. Otherwise, return the index in `v` of target where it would be if it were inserted in order.

The input `v` is guaranteed to be sorted and to contain unique values (i.e. no duplicates).

``` r
x <- c(1, 3, 5, 6)
search_insert_position(x, 5)
```

``` r        
## [1] 3
```

``` r
search_insert_position(x, 2)
```

``` r        
## [1] 2
```

``` r
search_insert_position(x, 7)
```

``` r        
## [1] 5
```

*Hint*: My solution considers various cases separately. I used the functions %in% that tests membership, which
to find TRUE indices of a logical vector, and all which tests if all entries of a logical vector are TRUE. Pay
particular attention to the final test case above.

**Testing membership with %in%:**

``` r
"cat" %in% c("dog", "cow", "cat", "owl")
```

``` r
## [1] TRUE
```

``` r
12 %in% c(3, 6, 1, 0)
```

``` r
## [1] FALSE
```

Using all:

``` r
all(c(T, T, T))
```

``` r
## [1] TRUE
```

```r
all(c(T, T, F, T))
```

```r
## [1] FALSE
```