---
title: "PSTAT 10 Worksheet 2"
subtitle: "Due 8/2/2022 11:59pm"
author: "Ingmar Sturm"
output: pdf_document
---

# Problem 1: Basic vector manipulation

1. Recall from lecture my 2021 monthly gas bill in order was given by:

```{r}
gasbill <- c(46, 33, 39, 37, 46, 30, 48, 32, 49, 35, 30, 48)
```

It turns out the charge for December should have been 49 instead of 48. Update the ```gasbill``` to reflect the
true charge. Try not to “cheat” and just type in all the old values again; use the existing ```gasbill``` vector.

2. Recreate the following numeric vector. Avoid typing in all of the values manually.

```{r, eval=FALSE}
## [1] -50 -51 -52 -53 -54 -53 -52 -51 -50
```

3. Create a vector from 1 to 10 with increments of 0.05. What is the length of this vector? Hint: Use `seq`
with `by` argument.

4. Create a vector of length 100 from 1 to 10 with uniform increments. What is the increment? Hint: Use
`seq` with `length` argument.

5. What happens if you try to use `seq` with both the `length` and `by` arguments specified?

# Problem 2

Download the file `ws2.csv` from the course website and import it into R. This data set has two variables
named `x` and `y`.

```{r}
library(here)
ws2_df <- read.csv(here("assignments", "week1_files", "ws2.csv"))
summary(ws2_df)
```

Remember the variables in a data frame are accessed by name with the dollar sign (and that the result is a
vector).

1. Determine the lengths of x and y.
2. What is the 40th element of x and the 80th element of y?
3. What is the average of all the values in the data frame, including both x and y?
4. How many elements of x are greater than 70?

Let’s look at the first 4 elements of x and y:

```{r}
ws2_df$x[1:4]
## [1] 74 89 78 23
ws2_df$y[1:4]
## [1] 58 26 48 80
```

The first three elements of x are greater than or equal to their corresponding 
element in y: $74 > 58$, $89 > 26$, $78 > 48$. But the fourth element of `x`, 
23, is less than the fourth element of `y`, 80.

5. How many elements of x are greater than or equal to the corresponding element in y?
6. What is the proportion of elements of x that are greater than or equal to the corresponding element in
y?
7. How many values in x differ from their corresponding value in y by more than 10 in absolute value?
Hint: there is an `abs` function.

# Problem 3

Create a vector of integers from 1 to 12 inclusive.

1. Use the vector to create a 3x4 matrix. Did recycling occur?
2. Use the vector to create a 4x4 matrix. Did recycling occur?

# Problem 4

Use the heights_df data frame from worksheet 1. The height variable is given in centimeters (cm).

1. Write a vectorized function cm_to_inch that takes a numeric centimeter and converts it to inches.

Apply the function to the height vector. First 10 elements are shown below:

```{r}
head(cm_to_inch(heights_df$height), 10) # the head function gives the first elements
## [1] 62.40 67.08 65.52 71.37 68.25 73.71 60.84 65.13 76.05 64.35
```

2. Write a vectorized function cm_to_ft_inch that converts numerical values given in cm to a feet inch
format, rounding to the nearest inch.
For example,

```{r}
cm_to_ft_inch(178)
## [1] "5 9"
```

You may need the (vectorized) quotient function %/% and the remainder function %%:

```{r}
# Quotient: 3 goes into 7 two times
7 %/% 3
## [1] 2
# Remainder: The remainder when 7 is divided by 3 is one
7 %% 3
## [1] 1
```

Remember you should look things up on StackOverflow if you’re stuck with some operations.
Apply the function to the height vector.

```{r}
head(cm_to_ft_inch(heights_df$height), 10)
## [1] "5 2" "5 7" "5 6" "5 11" "5 8" "6 2" "5 1" "5 5" "6 4" "5 4"
```