---
title: "PSTAT 10 Worksheet 2 Solutions"
output: html_document
---

## Problem 1: Basic vector manipulation

1.
```{r}
gasbill <- c(46, 33, 39, 37, 46, 30, 48, 32, 49, 35, 30, 48)
gasbill[12] <- 49
gasbill
```

2.
```{r}
c(-50:-54, -53:-50)
```

3.
```{r}
x <- seq(1, 10, by = 0.05)
length(x)
```

4.
```{r}
x <- seq(1, 10, length = 100)
length(x)
```

5.
You should get an error that says “too many arguments”. You are constrained to specify at most one of by and length, which makes sense.

## Problem 2

```{r}
library(here)
ws2_df <- read.csv(here("week1_files", "ws2.csv"))
summary(ws2_df)
```

Determine the lengths of `x` and `y`.

```{r}
c(length(ws2_df$x), length(ws2_df$y))
```
The lengths of variables in a data frame must be the same.

What is the 40th element of `x` and the 80th element of `y`?

```{r}
c(ws2_df$x[40], ws2_df$y[80])
```

What is the average of all the values in the data frame, including both `x` and `y`?

```{r}
mean(c(ws2_df$x, ws2_df$y))
```

How many elements of `x` are greater than 70?

```{r}
sum(ws2_df$x > 70)
```

How many elements of `x` are greater than or equal to the corresponding element in `y`?

```{r}
sum(ws2_df$x >= ws2_df$y)
```

What is the proportion of elements of `x` that are greater than or equal to the corresponding element in `y`?

```{r}
mean(ws2_df$x >= ws2_df$y)
```

Why does the above give the proportion? Because the average of the sum of the components of a binary vector is the proportion of 1s. We will use this trick later when we discuss simulation.

How many values in `x` differ from their corresponding value in `y` by more than 10 in absolute value?

```{r}
sum(abs(ws2_df$x - ws2_df$y) > 10)
```
You should work out why the above gives the right answer. Vectorization is very handy for these kinds of operations.

## Problem 3

Create a vector of integers from 1 to 12 inclusive.

```{r}
x <- 1:12
```

Use the vector to create a 3x4 matrix. Did recycling occur?

```{r}
mx <- matrix(x, nrow = 3, ncol = 4)
mx
```
Recycling did not occur; all elements of `x` were used exactly once.

Use the vector to create a 4x4 matrix. Did recycling occur?

```{r}
mx <- matrix(x, nrow = 4, ncol = 4)

mx
```
Recycling occurred; the last column was filled with the first column. We got an error message since the size of the matrix is not a multiple of 12.

## Problem 4

```{r}
heights_df <- read.csv(here("week1_files", "heights.csv"))
```

Write a vectorized function `cm_to_inch` that takes a numeric centimeter and converts it to inches. Apply the function to the height vector.
1cm is approximately 0.39in.

```{r}
cm_to_inch <- function(cm) {
  cm * 0.39 # Remember the final expression in a function is automatically returned
}
```

```{r}
head(cm_to_inch(heights_df$height), 10)
```

Write a vectorized function `cm_to_ft_inch` that converts numerical values given in cm to a feet inch format, rounding to the nearest inch.

```{r}
cm_to_ft_inch <- function(cm) {
  inch <- cm_to_inch(cm)
  ft <- inch %/% 12
  inch <- round(inch %% 12)
  paste(ft, inch)
}
```

```{r}
head(cm_to_ft_inch(heights_df$height), 10)
```