---
title: PSTAT 10 Worksheet 3 Solutions
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

## Problem 1: Contains Duplicate

Write the function `contains_duplicate(v)` that takes a numeric vector `v` and returns TRUE if any value appears **at least twice** in the vector and FALSE otherwise.

**There are several ways to do this**. The *extremely* easy way is it use the `duplicated` function.

```{r}
contains_duplicate <- function(v) {
  any(duplicated(v))
}
contains_duplicate(c(1, 2, 3, 1))
```

```         
## [1] TRUE
```

```{r}
contains_duplicate(c(1, 2, 3, 4))
```

```         
## [1] FALSE
```

```{r}
contains_duplicate(c(1, 1, 1, 3, 3, 4, 3, 2, 4, 2))
```

```         
## [1] TRUE
```

Another way involves using a loop:

```{r}
contains_duplicate <- function(v) {
  seen <- rep(NA, length(v)) # Initialize with NA values
  for (i in seq_along(v)) {
    if (v[i] %in% seen) {
      return(TRUE)
    }
    seen[i] <- v[i]
  }
  return(FALSE)
}
```

We haven’t talked about `NA` yet, but we need it above for a technical reason: initializing with `vector(length = length(v))` would create a vector containing zeros. This leads to wrong output if `v` contain a zero.

## Problem 2: More on iris

1.  Convert the `iris` data frame to a tibble and call it `iris_tbl`

```{r, include=F}
library(tidyverse)
data("iris")
```

```{r}
iris_tbl <- as_tibble(iris)
```

1.  Find the median `Petal.Width` and then create a tibble that only contains petal widths greater than the median.

```{r}
median(iris_tbl$Petal.Width)
```


```{r}
iris_tbl |>
  filter(Petal.Width > median(iris_tbl$Petal.Width))
```

1.  Call the area of a petal its length times its width. Create a tibble containing only the variables `Sepal.Length`, `Sepal.Width`, `Species`, and `Petal.Area` and only the rows where the petal width is greater than the median.

```{r}
iris_tbl |>
  filter(Petal.Width > median(Petal.Width)) |>
  mutate(Petal.Area = Petal.Width * Petal.Length) |>
  select(-Petal.Length, -Petal.Width)
```

## Problem 3: More on heights data

Load the `heights_df` data frame from worksheet 1.

```{r include=FALSE}
heights_df <- read_csv("~/Documents/GitHub/essentialols.github.io/courses/pstat10_2024/assignments/week1_files/heights.csv")
```

```{r eval=FALSE}
library(here)
heights_df <- read.csv(here("heights.csv"))
```

Recall the `height` variable is given in centimeters (cm). In worksheet 2, we created `cm_to_ft_inch` that converts from cm to a string representation of feet and inches.

Using `dplyr` functionality, create a tibble with a variable `height_ft_in` in place of `height`.

```{r include=FALSE}
cm_to_ft_inch <- function(height){
  total_height_inches <- height/2.54
  height_feet <- round(total_height_inches/12)
  height_inches <- round(total_height_inches %% 12)
  return(paste(height_feet, height_inches))
}
```

```{r}
heights_df |>
  as_tibble() |>
  mutate(height_ft_in = cm_to_ft_inch(height)) |>
  select(-height)
```
