# Lecture 3 demo code

#### 2024-06-26

### Functions (slide 7)

1.  

```{r}
increment_power <- function(x, pwr = 2) {
  x <- x + 1
  return(x^pwr)
}
```

2.  

```{r}
increment_power(1:4, pwr = 2)
```

The vector `2` was recycled to a length four vector. The vector `1` was
recycled to a length four vector and each component of `1:4` is
incremented. Then a vectorized power function was applied. The above is
equivalent to

```{r}
(2:5)^c(2, 2, 2, 2)
```

This is vectorized because `^` is a vectorized operator; the result is a
vector where each component of the base is raised to the corresponding
component of the exponent.

```{r}
increment_power(2:5, pwr = c(2, 3))
```

Similar situation: the shorter vector `c(2, 3)` is recycled to be length
4. The above is equivalent to

```{r}
(2:5)^c(2, 3, 2, 3)
```

### Branching (slide 13)

```{r}
noise <- function(animal) {
  if (animal == "cow") {
    return("moo")
  }
  else if (animal == "dog") {
    return("woof")
  }
  else if (animal == "pig") {
    return("oink")
  }
  else if (animal == "owl") {
    return("hoot")
  } else {
    return("Animal not recognized!")
  }
}

noise("pig")
```

```{r}
noise("hedgehog")
```

```{r}
noise("cow")
```

### Branching (slide 14)

```{r}
parity <- function(x) {
  if (x %% 2 == 0) {
    return("even")
  } else if (x %% 2 == 1) {
    return("odd")
  } else {
    return("Not an integer!")
  }
}
parity(4)
```

```{r}
parity(17)
```

```{r}
parity(12.43)
```

### Function with branch (slide 19)

```{r}
parity_vec <- function(x) {
  # result <- vector(length = length(x)) # initialize result
  result <- NULL
  for (i in seq_along(x)) {
    result[i] <- parity(x[i]) # call the previously defined function
  }
  result # last expression is automatically returned
}

parity_vec(c(12, 320, 598, 23))
```

We used `seq_along(x)` so that we have an index `i` to use with
`result`.

### Vectorized ifelse (slide 20)

```{r}
parity_vec <- function(x) {
  ifelse(x %% 2, "odd", "even")
}
parity_vec(c(12, 320, 598, 23))
```

This takes advantage of the fact that numeric `0` is logically FALSE and
`1` is logically TRUE. Actually, all numeric values other than `0` are
TRUE. In view of this, the above is actually incorrect; we need to
account for the `"Not an integer!"` case.

```{r}
parity_vec <- function(x) {
  result <- ifelse(x %% 2, "odd", "even")
  not_int <- (x %% 1) != 0  # This is determines if x contains decimals (not int)
  result[not_int] <- "Not an integer!"
  result
}
parity_vec(c(12, 320.5, 599, 23.88))
```
