# Lecture 5 demo code 

#### Ingmar

#### 2024-07-01

We need to install and then load the following packages

```{r}
library(tidyverse)
```

```{r}
library(gapminder)
library(datasets)
```

```{r}
library(gapminder)

head(gapminder, 15)
```
```{r}
tail(gapminder, 15)
```

## Basic Plotting (slide 32)

```{r}
hist(gapminder$lifeExp, main = "Life Expectancy", xlab = "years")
```

## Plotting variables against each other (slide 35)

```{r}
plot(lifeExp ~ gdpPercap, gapminder, main = "Life Expectancy vs. GDP",
     xlab = "GDP per capita", ylab = "Life Expectancy")
```

Instead of specifying the second argument, we could plot the vectors
directly:

```{r}
plot(gapminder$lifeExp ~ gapminder$gdpPercap) # no second argument
```

The above plot is clearer on the *log-scale*. This is because countries
tend to be *exponentially* wealthier than each other in terms of GDP per
capita. Note for instance that the x-axis in the above plot contains
large values.

```{r}
plot(lifeExp ~ log(gdpPercap), gapminder,
     main = "Life Expectancy and GDP on log scale")
```

The relationship is much clearer after transforming one of the
variables!
