---
title: "Lecture 8 demo code"
date: 2024-07-08
output: html_document
---

## Setting a seed

The following ensures you will get the same results as me.

```{r}
set.seed(10)
```


## Minimum of Scores Greater than 2 (slide 13)

```{r}
min_greater_2 <- function() {
  s <- sample(1:6, 3, replace = T)
  return(min(s) > 2)
}

r <- replicate(10000, min_greater_2())
mean(r)
```


The following is a little obnoxious, but we can combine it all into a single line.

```{r}
replicate(10000, min(sample(1:6, 3, replace = T)) > 2) |>
  mean()
```

## More dice rolling (slide 14)

```{r}
second_greater <- function() {
  s <- sample(1:6, 2, replace = T)
  s[2] > s[1]
}
mean(replicate(10000, second_greater()))
```

The `diff` function gives the lagged difference; the difference between successive elements. Below, the difference between 4 and 0 is 4 and the distance between 10 and 4 is 6.

```{r}
diff(c(0, 4, 10))
```

We can use this to construct a one-liner:

```{r}
replicate(10000, diff(sample(1:6, 2, replace = T)) > 0) |> 
  mean()
```

I don’t expect you to remember the diff function; just know that many basic operations are already available for you in base R.

## Derangement

Below is a solution using loops.

```{r}
x <- 1:100
is_deranged_loop <- function(){
  counta <- 0
  x1 <- sample(x)
  for(i in seq_along(x)){
    if(x1[i] == i){
      counta = counta + 1
    }
  }
  return(counta <= 0)
}
mean(replicate(10000, is_deranged_loop()))
```

And here is a vectorized solution.

```{r}
is_deranged <- function() {
 !any(sample(1:100) == 1:100) # Think about what this does!
}
result <- replicate(10000, is_deranged())
mean(result)
```

## Plotting the approximation

The function `running_mean(k)` gives the average of the first k elements of result, i.e. the estimate from the first k replications.

```{r}
running_mean <- function(k) {
  mean(result[seq_len(k)])
}

running_means <- sapply(1:10000, running_mean)

head(running_means)
```

Actually, there is a much smarter way to do the above using the cumsum function. This uses a vectorized division.

```{r}
running_means <- cumsum(result) / 1:10000
```

Plot these estimates against the replication number.

```{r}
plot(1:10000, running_means, type = 'l', main = "Probability of a derangement",
     xlab = "no. reps", ylab = "p")
abline(h = 1/exp(1), col = "red")
```

## Expectation (slide 20)

```{r}
set.seed(100)
before_six <- function() {
  s <- sample(1:6, 100, replace= T)
  which(s == 6)[1] - 1 # Minus 1 since we do not count the roll which scored 6.
}
replicate(10000, before_six()) |> mean()
```

It can be shown the theoretical value is 5.